---
title: "Cleaning biomass data - individual count and dry mass"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(janitor)
library(skimr)
library(purrr)
library(tidyr)

source(here("R/func/clean_functions.R"))
```

# Cleaning biomass data

Reading in data and creating observation number

```{r, warning=FALSE}
biomass_data_taxaclean <-
  read_csv(here("data/processed/biomass_data_cleannames.csv")) |> 
  mutate(obs_num = row_number())
```

## Pre-cleaning the data

Setting a minimum mass cutoff (due to higher proportional measurement error at very small mass values). This also acts to drop NA's for individual count when included instead of a mass measure. 

```{r}
step_1 <- min_cutoff(data = biomass_data_taxaclean, individual_count, dm_g) 
```

Filtering formalin samples only

```{r}
step_2 <- step_1 |>
  filter(preservation_method == "Formalin")
```

Filtering taxa that have enough observations for analysis (at least 10)

```{r}
step_3 <- taxa_enough_obs(step_2, taxa_clean)

# checking number of observations of included taxa
step_3 |>
   pull(taxa_clean) |> tabyl()
```

## Identifying outlier candidates

### Cleveland Plots

Creating Cleveland plots for each taxa as a visual tool to help identify observations that are far-removed from the rest

```{r}
cleveland_plots <- step_3 |>
  group_nest(taxa_clean) |>
  mutate(ind_count_plot = map2(.x = data,
                    .y = taxa_clean,
                    ~cleveland_plot(.x, "individual_count") + ggtitle(.y)),
         dm_plot = map2(.x = data,
                    .y = taxa_clean,
                    ~cleveland_plot(.x, "dm_g") + ggtitle(.y)))
cleveland_plots$ind_count_plot
cleveland_plots$dm_plot
```

### Percentage analysis 

Identifying top and bottom 2.5% of the data for the mass or count variables being analysed

```{r}
biomass_formalin_percentiles <- step_3 |>
  group_by(taxa_clean) |>
  mutate(keep_data_ind_count_percentiles = percentiles(individual_count, 0.025, 0.975),
         keep_data_dm_percentiles = percentiles(dm_g, 0.025, 0.975))
  
# Checking which observations are identified by the percentile method
biomass_formalin_percentiles |> 
  filter(keep_data_ind_count_percentiles == FALSE) |> 
  select(taxa_clean, individual_count, keep_data_ind_count_percentiles, obs_num) |>
  arrange(taxa_clean, individual_count) |>
  print(n = Inf)
```

### Measures of outlyingness

Creating logical vectors identifying which observations fall outside the cutoff value or range for each measure of outlyingness (cutoffs calculated according to Aguinis, H., R. K. Gottfredson and H. Joo. Best-Practice Recommendations for Defining, Identifying, and Handling Outliers. Organizational Research Methods, 16, 270-301 (2013))

```{r}
#Nesting by taxa
taxa_nested <- biomass_formalin_percentiles |> 
  select(taxa_clean, individual_count, dm_g, obs_num, keep_data_ind_count_percentiles, keep_data_dm_percentiles) |> 
  nest(data = c(individual_count, dm_g, obs_num, keep_data_ind_count_percentiles, keep_data_dm_percentiles))
  
#Calculating measures of outlyingness and associated cutoff values
taxa_nested_outlier_candidates_identified <- taxa_nested |> 
  mutate(keep_data_leverage = map(.x = data,
                                  ~ keep_data_leverage(.x, "individual_count", "dm_g")),
         keep_data_stud_resid = map(.x = data,
                                    ~ keep_data_stud_resid(.x, "individual_count", "dm_g")),
         keep_data_cook = map(.x = data,
                              ~ keep_data_cook(.x, "individual_count", "dm_g")),
         keep_data_dffits = map(.x = data,
                                ~ keep_data_dffits(.x, "individual_count", "dm_g")),
         keep_data_dfbetas = map(.x = data,
                                ~ keep_data_dfbetas(.x, "individual_count", "dm_g")))
```

Checking which observations of a particular taxa are identified by different measures of outlyingess

```{r}
taxa_unnested_outlier_candidates_identified <- taxa_nested_outlier_candidates_identified |>
  unnest(cols = c(data, keep_data_leverage, keep_data_stud_resid, keep_data_cook, 
                  keep_data_dffits, keep_data_dfbetas)) 

taxa_unnested_outlier_candidates_identified |>
  filter(taxa_clean == "Amphipoda") |>
  select(keep_data_leverage, obs_num) |>
  filter(keep_data_leverage == FALSE) |>
  print(n = Inf)
```

Consolidating all outlier candidates into one column

```{r}
taxa_unnested_keep_data_merged <- taxa_unnested_outlier_candidates_identified |>
 # assigning dfbetas values for intercept and slope to separate vectors
   mutate(keep_data_dfbetas_int = keep_data_dfbetas[,1], 
         keep_data_dfbetas_slope = keep_data_dfbetas[,2]) |>
 # Consolidating all outlier candidates into one vector
   mutate(keep_data = ifelse(keep_data_ind_count_percentiles == TRUE & keep_data_dm_percentiles == TRUE & keep_data_cook == TRUE & keep_data_leverage == TRUE & keep_data_stud_resid == TRUE & keep_data_dffits == TRUE & keep_data_dfbetas_int == TRUE & keep_data_dfbetas_slope == TRUE, TRUE, FALSE)) 
```

### Summary table of counts for all outlier methods 

```{r}
long_data <- taxa_unnested_keep_data_merged |> 
  ungroup() |> 
  select(-c(keep_data_dfbetas)) |> 
  pivot_longer(starts_with("keep"), names_to = "outlier_method") 

long_data |> 
  group_by(outlier_method) |> 
  summarise(n = n(),
            n_FALSE = length(which(value == FALSE)),
            n_TRUE = length(which(value == TRUE)),
            n_NA = length(which(is.na(value))),
            n_add_check = n_FALSE + n_TRUE + n_NA,
            percent_false_or_NA = (n_FALSE+n_NA)/(n_FALSE+n_TRUE+n_NA)*100
  )
```

## Saving the data
```{r}
biomass_formalin_count_dm_cleaned <- taxa_unnested_keep_data_merged |>
  select(taxa_clean, individual_count, dm_g, keep_data)

write_csv(biomass_formalin_count_dm_cleaned,
          here("data/processed/formalin/count_dm/biomass_formalin_count_dm_cleaned.csv"))
```

