---
title: "R Notebook"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(smatr)
library(janitor)
library(skimr)
library(gridExtra)
library(olsrr)
library(purrr)
library(tidyr)

source(here("R/func/clean_functions.R"))
```

# Read in data and create observation numer

```{r, warning=FALSE}
biomass_data_taxaclean <-
  read_csv(here("data/processed/biomass_data_cleannames.csv")) |> 
  mutate(obs_num = row_number())
```

## Pre-cleaning the data

setting minimum mass cut-off (due to higher proportional measurement error at very small mass values) and removing observations when wet mass is less than dry mass (considering these errors in measurement or recording)
```{r}
biomass_data_taxaclean_mass_cutoff <- biomass_data_taxaclean |>
   filter(wm_g >= 0.001, dm_g >=0.001, wm_g >= dm_g)
 
  # filter(wm_g >= 0.001, dm_g >=0.001, afdm_g >=0.001, wm_g >= dm_g, dm_g >= afdm_g) |>
  # select(source, taxa_clean, preservation_method, wm_g, dm_g, afdm_g) |>
  # print(n = Inf)

```

```{r}
step_1 <- min_cutoff(data = biomass_data_taxaclean, wm_g, dm_g)  
step_2 <- check_mass_diff(step_1,  wm_g, dm_g)
```

filtering frozen samples only
```{r}
step_3 <- step_2 |>
  filter(preservation_method == "Frozen")
```

filter taxa that have enough obs for analysis

find taxa less than 10 obs

```{r}
notenough_taxa <- step_3 |> 
  group_by(taxa_clean) |> 
  summarise(n = n()) |> 
  filter(n < 10) |> 
  pull(taxa_clean)

# Checking
step_3 |> 
    filter(! taxa_clean %in% notenough_taxa) |> 
  pull(taxa_clean) |> tabyl()
```


```{r}
taxa_notenough_obs(step_3, taxa_clean) |> 
  pull(taxa_clean) |> tabyl()

step_4 <- taxa_notenough_obs(step_3, taxa_clean)
```

## Identifying outlier candidates

### Cleveland Plots
Creating Cleveland plots for each taxa as a visual tool to help identify observations that are far-removed from the rest.
```{r fig.height=15, fig.width=10}

cleveland_plots <- step_4 |>
  group_nest(taxa_clean) |>
  mutate(plot = map2(.x = data,
                    .y = taxa_clean,
                    ~cleveland_plot(.x, "wm_g") + ggtitle(.y)))
cleveland_plots$plot

```

### Percentage analysis 
Identifying top and bottom 2.5% of the data for each mass variable
```{r}

#biomass_frozen_percentiles <- biomass_wobs |>
#  mutate(keep_data_wm_percentiles = ifelse(wm_g >= quantile(wm_g, 0.025, na.rm = TRUE) & wm_g <= quantile(wm_g, #0.975, na.rm = TRUE), TRUE, FALSE),
#        keep_data_dm_percentiles = ifelse(dm_g >= quantile(dm_g, 0.025, na.rm = TRUE) & dm_g <= quantile(dm_g, #0.975, na.rm = TRUE), TRUE, FALSE))
  
biomass_frozen_percentiles <- step_4 |>
  group_by(taxa_clean) |>
  mutate(keep_data_mass1_percentiles = biomass_percentiles(wm_g, 0.025, 0.975),
         keep_data_mass2_percentiles = biomass_percentiles(dm_g, 0.025, 0.975))
  
# Checking which observations are identified by the percentile method
biomass_frozen_percentiles |> 
  filter(keep_data_mass1_percentiles == FALSE) |> 
  select(taxa_clean, wm_g, keep_data_mass1_percentiles, obs_num) |>
  arrange(taxa_clean, wm_g) |>
  print(n = Inf)
```

### Measures of outlyingness
Creating logical vectors identifying which observations fall outside the cutoff value or range for each measure of outlyingness
```{r}

#Nesting by taxa
taxa_nested <- biomass_frozen_percentiles |> 
  select(taxa_clean, wm_g, dm_g, obs_num, keep_data_mass1_percentiles, keep_data_mass2_percentiles) |> 
  nest(data = c(wm_g, dm_g, obs_num, keep_data_mass1_percentiles, keep_data_mass2_percentiles))
  
#Calculating measures of outlyingness and associated cutoff values
taxa_nested_outlier_candidates_identified <- taxa_nested |> 
  mutate(keep_data_leverage = map(.x = data,
                                  ~ keep_data_leverage(.x, "wm_g", "dm_g")),
         keep_data_stud_resid = map(.x = data,
                                    ~ keep_data_stud_resid(.x, "wm_g", "dm_g")),
         keep_data_cook = map(.x = data,
                              ~ keep_data_cook(.x, "wm_g", "dm_g")),
         keep_data_dffits = map(.x = data,
                                ~ keep_data_dffits(.x, "wm_g", "dm_g")),
         keep_data_dfbetas = map(.x = data,
                                ~ keep_data_dfbetas(.x, "wm_g", "dm_g")))

```

Checking which observations of a particular taxa are identified by different measures of outlyingess
```{r}

taxa_unnested_outlier_candidates_identified <- taxa_nested_outlier_candidates_identified |>
  unnest(cols = c(data, keep_data_leverage, keep_data_stud_resid, keep_data_cook, 
                  keep_data_dffits, keep_data_dfbetas)) 

taxa_unnested_outlier_candidates_identified |>
  filter(taxa_clean == "Amphipoda") |>
  select(keep_data_leverage, obs_num) |>
  filter(keep_data_leverage == FALSE) |>
  print(n = Inf)

```

Consolidating all outlier candidates into one vector
```{r}

taxa_unnested_keep_data_merged <- taxa_unnested_outlier_candidates_identified |>
 # assigning dfbetas values for intercept and slope to separate vectors
   mutate(keep_data_dfbetas_int = keep_data_dfbetas[,1], 
         keep_data_dfbetas_slope = keep_data_dfbetas[,2]) |>
 # Consolidating all outlier candidates into one vector
   mutate(keep_data = ifelse(keep_data_mass1_percentiles == TRUE & keep_data_mass2_percentiles == TRUE & keep_data_cook == TRUE & keep_data_leverage == TRUE & keep_data_stud_resid == TRUE & keep_data_dffits == TRUE & keep_data_dfbetas_int == TRUE & keep_data_dfbetas_slope == TRUE, TRUE, FALSE)) 

```

### Getting a summary table of counts for all outlier methods 
```{r}

long_data <- taxa_unnested_keep_data_merged |> 
  ungroup() |> 
  select(-c(keep_data_dfbetas)) |> 
  pivot_longer(starts_with("keep"), names_to = "outlier_method") 

long_data |> 
  group_by(outlier_method) |> 
  summarise(n = n(),
            n_FALSE = length(which(value == FALSE)),
            n_TRUE = length(which(value == TRUE)),
            n_NA = length(which(is.na(value))),
            n_add_check = n_FALSE + n_TRUE + n_NA,
            percent_false_or_NA = (n_FALSE+n_NA)/(n_FALSE+n_TRUE+n_NA)*100
  )

```

# Save the data
```{r}
biomass_frozen_outliers_identified <- taxa_unnested_keep_data_merged |>
  select(taxa_clean, wm_g, dm_g, keep_data)

write_csv(biomass_frozen_outliers_identified, here("data/processed/biomass_frozen_outliers_identified.csv"))
```

