---
title: "Regression analysis"
author: "Josh Nitschke"
date: "2023-08-31"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(janitor)
library(skimr)
library(purrr)
library(tidyr)
```

## Reading in cleaned data

```{r}
biomass_ethanol_dm_afdm_cleaned <- read_csv(here("data/processed/ethanol/dm_afdm/biomass_ethanol_dm_afdm_cleaned.csv"))
```

## Running lm's on untransformed data

```{r}
taxa_nested_lm <- biomass_ethanol_dm_afdm_cleaned |> 
  group_nest(taxa_clean) |>
  mutate(lm_object = map(.x = data,
                          ~ lm(afdm_g ~ dm_g, data = .x)))
```

### Checking assumptions

#### Homogeneity of variance

```{r}
taxa_nested_lm_fitted_resid <- taxa_nested_lm |> 
  mutate(fitted_resid = map(.x = lm_object,
                             .f = ~ tibble(fitted = fitted(.x), residuals = residuals(.x))),
         fitted_resid_plot = map2(.x = fitted_resid,
                     .y = taxa_clean,
                     .f = ~ ggplot(data = .x, aes(x = fitted, y = residuals)) + 
                       ggtitle(.y) +
                       labs(x = "fitted value", y = "residual") +
                       geom_point() + 
                       geom_hline(yintercept = 0, linetype = 2) +
                       theme_bw() + 
                       theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
                       panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), 
                       axis.text = element_text(size = 12, colour = "black"), axis.title = element_text(size = 14))))
taxa_nested_lm_fitted_resid$fitted_resid_plot
```

#### Normality

```{r}
taxa_nested_lm_fitted_resid_qq <- taxa_nested_lm_fitted_resid |>
  mutate(qq = map2(.x = lm_object,
                  .y = taxa_clean,
                  .f = ~ ggplot(.x, aes(sample = residuals(.x))) + 
                    geom_qq(aes(sample = residuals(.x))) + 
                    stat_qq_line(aes(sample = residuals(.x))) +
                    ggtitle(.y) +
                    labs(x = "theoretical quantiles", y = "sample quantiles") +
                    theme_bw() + 
                    theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
                    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), 
                    axis.text = element_text(size = 12, colour = "black"), axis.title = element_text(size = 14),
                    title = element_text(size = 18))))
taxa_nested_lm_fitted_resid_qq$qq
```

## Excluding outliers

### Running lm's on untransformed data

```{r}
taxa_nested_lm_outliers_removed <- biomass_ethanol_dm_afdm_cleaned |> 
  filter(keep_data == TRUE) |>
  group_nest(taxa_clean) |>
  mutate(lm_object = map(.x = data,
                          ~ lm(afdm_g ~ dm_g, data = .x)))
```

### Checking assumptions

#### Homogeneity of variance

```{r}
taxa_nested_lm_outliers_removed_fitted_resid <- taxa_nested_lm_outliers_removed |> 
  mutate(fitted_resid = map(.x = lm_object,
                             .f = ~ tibble(fitted = fitted(.x), residuals = residuals(.x))),
         fitted_resid_plot = map2(.x = fitted_resid,
                     .y = taxa_clean,
                     .f = ~ ggplot(data = .x, aes(x = fitted, y = residuals)) + 
                       ggtitle(.y) +
                       labs(x = "fitted value", y = "residual") +
                       geom_point() + 
                       geom_hline(yintercept = 0, linetype = 2) +
                       theme_bw() + 
                       theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
                       panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), 
                       axis.text = element_text(size = 12, colour = "black"), axis.title = element_text(size = 14))))
taxa_nested_lm_outliers_removed_fitted_resid$fitted_resid_plot
```

#### Normality

```{r}
taxa_nested_lm_outliers_removed_fitted_resid_qq <- taxa_nested_lm_outliers_removed_fitted_resid |>
  mutate(qq = map2(.x = lm_object,
                  .y = taxa_clean,
                  .f = ~ ggplot(.x, aes(sample = residuals(.x))) + 
                    geom_qq(aes(sample = residuals(.x))) + 
                    stat_qq_line(aes(sample = residuals(.x))) +
                    ggtitle(.y) +
                    labs(x = "theoretical quantiles", y = "sample quantiles") +
                    theme_bw() + 
                    theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
                    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), 
                    axis.text = element_text(size = 12, colour = "black"), axis.title = element_text(size = 14),
                    title = element_text(size = 18))))
taxa_nested_lm_outliers_removed_fitted_resid_qq$qq
```

## Log-transforming to stabilise variance and improve normality

### Running lm's on ln-transformed data and plotting

```{r}
taxa_nested_lm_ln_plot <- biomass_ethanol_dm_afdm_cleaned |> 
  group_nest(taxa_clean) |>
  mutate(lm_ln_object = map(.x = data,
                            .f = ~ lm(log(afdm_g) ~ log(dm_g), data = .x)),
         lm_ln_plots = map2(.x = data,
                         .y = taxa_clean,
                         .f = ~ ggplot(.x, aes(x = log(dm_g), y = log(afdm_g))) + 
                           geom_point() +
                           geom_smooth(method = "lm") +
                           labs(x = "log(e) dry mass (g)", y = "log(e) ash-free dry mass (g)") +
                           ggtitle(.y)))
           
taxa_nested_lm_ln_plot$lm_ln_plots
```

### Checking assumptions again

#### Homogeneity of variance

```{r}
taxa_nested_lm_ln_plot_fitted_resid <- taxa_nested_lm_ln_plot |> 
  mutate(fitted_resid = map(.x = lm_ln_object,
                             .f = ~ tibble(fitted = fitted(.x), residuals = residuals(.x))),
         fitted_resid_plot = map2(.x = fitted_resid,
                     .y = taxa_clean,
                     .f = ~ ggplot(data = .x, aes(x = fitted, y = residuals)) + 
                       ggtitle(.y) +
                       labs(x = "fitted value", y = "residual") +
                       geom_point() + 
                       geom_hline(yintercept = 0, linetype = 2) +
                       theme_bw() + 
                       theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
                       panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), 
                       axis.text = element_text(size = 12, colour = "black"), axis.title = element_text(size = 14))))
taxa_nested_lm_ln_plot_fitted_resid$fitted_resid_plot
```

#### Normality

```{r}
taxa_nested_lm_ln_plot_fitted_resid_qq <- taxa_nested_lm_ln_plot_fitted_resid |>
  mutate(qq = map2(.x = lm_ln_object,
                  .y = taxa_clean,
                  .f = ~ ggplot(.x, aes(sample = residuals(.x))) + 
                    geom_qq(aes(sample = residuals(.x))) + 
                    stat_qq_line(aes(sample = residuals(.x))) +
                    ggtitle(.y) +
                    labs(x = "theoretical quantiles", y = "sample quantiles") +
                    theme_bw() + 
                    theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
                    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), 
                    axis.text = element_text(size = 12, colour = "black"), axis.title = element_text(size = 14),
                    title = element_text(size = 18))))
taxa_nested_lm_ln_plot_fitted_resid_qq$qq
```

### Regression summary statistics

Pulling out regression summary statistics

```{r}
# Getting coefficients from lm
taxa_nested_lm_ln_plot_coef <- taxa_nested_lm_ln_plot |>
  mutate(summary = map(.x = lm_ln_object,
                       .f = ~ summary(.x)))
long_summary <- taxa_nested_lm_ln_plot_coef |>
  mutate(coefs = map(.x = summary,
                            ~.x[[4]]),
         r2 = map(.x = summary,
                            ~.x[[9]]),
         n = map(.x = lm_ln_object,
                 ~length(.x$residuals)),
         conf_interval = map(.x = lm_ln_object,
                             .f = ~confint(.x))) |>
  unnest(c(coefs, r2, n, conf_interval)) |>
  mutate(coef_type = c(rep(c("intercept", "slope"), times = length(taxa_clean)/2))) |>
  pivot_wider(names_from = coef_type, values_from = c(coefs, conf_interval))

intercept <- long_summary$coefs_intercept[,1]
intercept_se <- long_summary$coefs_intercept[,2]
intercept_ci_lower <- long_summary$conf_interval_intercept[,1]
intercept_ci_upper <- long_summary$conf_interval_intercept[,2]
slope <- long_summary$coefs_slope[,1]
slope_se <- long_summary$coefs_slope[,2]
slope_ci_lower <- long_summary$conf_interval_slope[,1]
slope_ci_upper <- long_summary$conf_interval_slope[,2]
p <- long_summary$coefs_slope[,4] # p-value of the slope is the same as the p-vale of the overall regression model in the case of simple regression with one predictor
long_summary_short <- long_summary |>
  select(taxa_clean, n, r2)
  
reg_summary <- add_column(long_summary_short, tibble(p, slope, slope_se, slope_ci_lower, slope_ci_upper, intercept, intercept_se, intercept_ci_lower, intercept_ci_upper))

# Back-transforming intercept to get parameter 'a' in power equation: y = ax^b
reg_summary_a <- reg_summary |> 
  mutate(a = exp(intercept),
         a_ci_lower = exp(intercept_ci_lower),
         a_ci_upper = exp(intercept_ci_upper))
```

Calculating SF (smearing factor - mean of back-transformed residuals) as per Mährlein, M., M. Pätzig, M. Brauns and A. M. Dolman. Length–mass relationships for lake macroinvertebrates corrected for back-transformation and preservation effects. Hydrobiologia, 768, 37-50 (2016).

```{r}
taxa_nested_lm_ln_plot_coef_sf <- taxa_nested_lm_ln_plot_coef |>
  mutate(sf = map(.x = lm_ln_object,
                  .f = ~ (1/length(residuals(.x)))*sum(exp(residuals(.x)))))

taxa_sf <- taxa_nested_lm_ln_plot_coef_sf |>
  select(taxa_clean, sf) |>
  unnest(sf)

# Adding smearing factor to lm summary table
reg_summary_a_sf <- merge(reg_summary_a, taxa_sf)

# Saving the regression summary
write_csv(reg_summary_a_sf,
          here("output/ethanol/dm_afdm/biomass_regression_summary_ethanol_dm_afdm.csv"))
```

### Back-transformation 

Plotting back-transformed regression line with smearing factor correction and 95% confidence band

```{r}
# Extracting lower and upper 95% confidence bounds on the transformed scale
taxa_nested_lm_ln_plot_coef_CI <- taxa_nested_lm_ln_plot_coef_sf |> 
  mutate(conf_int = map(.x = lm_ln_object,
                        .f = ~ predict(.x, interval = "confidence"))) |>
  unnest(cols = c(data, sf, conf_int))

lower_CI <- taxa_nested_lm_ln_plot_coef_CI$conf_int[,2]
upper_CI <- taxa_nested_lm_ln_plot_coef_CI$conf_int[,3]

# Back-transforming confidence bounds and plotting with back-transformed regression line
taxa_nested_lm_ln_plot_coef_CI_BT <- taxa_nested_lm_ln_plot_coef_CI |>
  add_column(lower_CI, upper_CI) |>
  mutate(lower_CI_line = exp(lower_CI)*sf,
         upper_CI_line = exp(upper_CI)*sf) |>
  select(taxa_clean, dm_g, afdm_g, lower_CI_line, upper_CI_line, sf) |>
  nest(data = c(dm_g, afdm_g, lower_CI_line, upper_CI_line)) |>
  right_join(reg_summary_a) |>
  mutate(BT_plot = pmap(.l = list(data, slope, intercept, sf, taxa_clean, r2),
                        .f = ~ggplot(..1, aes(x = dm_g, y = afdm_g)) +
                          geom_point(fill = "grey", size = 3, alpha = 0.5, stroke = 1) +
                          geom_function(fun = function(x) (exp(..3))*(x^..2)*..4, 
                                        col = "darkgreen", linewidth  = 1) +
                          geom_ribbon(aes(ymin = lower_CI_line, ymax = upper_CI_line), fill = "darkgreen", 
                                      alpha = 0.2) +
                          ggtitle(..5) + 
                          labs(x = "dry mass (g)", y = "ash-free dry mass (g)") + 
                          geom_text(x = 0.25*max(..1$dm_g), y = 0.85*max(..1$afdm_g), 
                                    label = deparse(bquote(paste("A = ", 
                                                                 .(format(round(exp(..3), digits = 2), nsmall = 2)),
                                                                 "D" ^.(format(round(..2, digits = 2), nsmall = 2))*
                                                                   " \u00D7 ",
                                                                 .(format(round(..4, digits = 2), nsmall = 2)), 
                                                                 sep = ""))),
                                    parse = TRUE, size = 4) +
                          geom_text(x = 0.25*max(..1$dm_g), y = 0.75*max(..1$afdm_g), 
                                    label = deparse(bquote(paste("R"^2, " = ", 
                                                                 .(format(round(..6, digits = 2), nsmall = 2)), 
                                                                 sep = ""))),
                                    parse = TRUE, size = 4) +
                          theme_bw() + 
                          theme(panel.border = element_blank(), 
                                panel.grid.major = element_blank(),
                                panel.grid.minor = element_blank(), 
                                axis.line = element_line(colour = "black"),
                                axis.text = element_text(size = 12, colour = "black"), 
                                axis.title = element_text(size = 14), 
                                title = element_text(size = 16)) + 
                          scale_x_continuous(expand = expansion(mult = c(0, 0.1)), limits = c(0, NA)) + 
                          scale_y_continuous(expand = expansion(mult = c(0, 0.1)), limits = c(0, NA))))
taxa_nested_lm_ln_plot_coef_CI_BT$BT_plot
```

### Calculating biomass ratios and associated summary statistics

```{r}
biomass_ethanol_dm_afdm_ratios <- biomass_ethanol_dm_afdm_cleaned |>
  mutate(afdm_dm = afdm_g/dm_g) |>
  group_by(taxa_clean) |>
  summarise(n = n(),
            mean_afdm_dm = mean(afdm_dm),
            sd_afdm_dm = sd(afdm_dm),
            lower_ci_mean_afdm_dm = mean_afdm_dm-0.95*(sd_afdm_dm/sqrt(n)),
            upper_ci_mean_afdm_dm = mean_afdm_dm+0.95*(sd_afdm_dm/sqrt(n)),
            geom_mean = prod(afdm_dm)^(1/n),
            min_afdm_dm = min(afdm_dm),
            median_afdm_dm = median(afdm_dm),
            max_afdm_dm = max(afdm_dm),
            range_dm = paste0(format(round(min(dm_g), digits = 3), nsmall = 3), 
                              "-", format(round(max(dm_g), digits = 3), 
                                          nsmall = 3)))

# Saving the ratio summary
write_csv(biomass_ethanol_dm_afdm_ratios,
          here("output/ethanol/dm_afdm/biomass_ratio_summary_ethanol_dm_afdm.csv"))
```

### Assessing model performance

Calculating mean percentage error (MPE) and mean absolute percentage error (MAPE) for regression and ratio methods

```{r}
pred_mass_ethanol_dm_afdm_error_eval <- taxa_nested_lm_ln_plot_coef_CI_BT |>
  mutate(ols_pred_afdm = pmap(.l = list(data, a, slope, sf),
                            .f = ~ ..2*(..1$dm_g)^..3*..4),
         ols_pred_afdm_percent_error = map2(.x = data,
                                          .y = ols_pred_afdm,
                                          .f = ~-((.x$afdm_g-.y)/.x$afdm_g)*100),
         ratio_pred_afdm = map2(.x = data,
                              .y = biomass_ethanol_dm_afdm_ratios$mean_afdm_dm,
                              .f = ~ .y*.x$dm_g),
         ratio_pred_afdm_percent_error = map2(.x = data,
                                          .y = ratio_pred_afdm,
                                          .f = ~-((.x$afdm_g-.y)/.x$afdm_g)*100),
         ols_total_PE = map2(.x = data,
                             .y = ols_pred_afdm,
                             .f = ~ (sum(.y-.x$afdm_g)/sum(.x$afdm_g))*100),
         ols_MPE = map2(.x = ols_pred_afdm_percent_error,
                        .y = n,
                        .f = ~ sum(.x)/.y),
         ols_MAPE = map2(.x = ols_pred_afdm_percent_error,
                       .y = n,
                       .f = ~ sum(abs(.x))/.y),
         ratio_total_PE = map2(.x = data,
                             .y = ratio_pred_afdm,
                             .f = ~ (sum(.y-.x$afdm_g)/sum(.x$afdm_g))*100),
         ratio_MPE = map2(.x = ratio_pred_afdm_percent_error,
                        .y = n,
                        .f = ~ sum(.x)/.y),
         ratio_MAPE = map2(.x = ratio_pred_afdm_percent_error,
                       .y = n,
                       .f = ~ sum(abs(.x))/.y)) |> 
  select(taxa_clean, ols_total_PE, ols_MPE, ols_MAPE, ratio_total_PE, ratio_MPE, ratio_MAPE) |> 
  unnest(c(ols_total_PE, ols_MPE, ols_MAPE, ratio_total_PE, ratio_MPE, ratio_MAPE))
```

## Excluding outlier candidates

### Running lm's on ln-transformed data and plotting

```{r}
taxa_nested_outliers_removed_lm_ln_plot <- biomass_ethanol_dm_afdm_cleaned |> 
  filter(keep_data == TRUE) |>
  group_nest(taxa_clean) |>
  mutate(lm_ln_object = map(.x = data,
                          ~ lm(log(afdm_g) ~ log(dm_g), data = .x)),
         lm_ln_plots = map2(.x = data,
                         .y = taxa_clean,
                         .f = ~ ggplot(.x, aes(x = log(dm_g), y = log(afdm_g))) + 
                           geom_point() +
                           geom_smooth(method = "lm") +
                           labs(x = "log(e) dry mass (g)", y = "log(e) ash-free dry mass (g)") +
                           ggtitle(.y)))
           
taxa_nested_outliers_removed_lm_ln_plot$lm_ln_plots
```

### Checking assumptions again

#### Homogeneity of variance

```{r}
taxa_nested_outliers_removed_lm_ln_plot_fitted_resid <- taxa_nested_outliers_removed_lm_ln_plot |> 
  mutate(fitted_resid = map(.x = lm_ln_object,
                             .f = ~ tibble(fitted = fitted(.x), residuals = residuals(.x))),
         fitted_resid_plot = map2(.x = fitted_resid,
                     .y = taxa_clean,
                     .f = ~ ggplot(data = .x, aes(x = fitted, y = residuals)) + 
                       ggtitle(.y) +
                       labs(x = "fitted value", y = "residual") +
                       geom_point() + 
                       geom_hline(yintercept = 0, linetype = 2) +
                       theme_bw() + 
                       theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
                       panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), 
                       axis.text = element_text(size = 12, colour = "black"), axis.title = element_text(size = 14))))
taxa_nested_outliers_removed_lm_ln_plot_fitted_resid$fitted_resid_plot
```

#### Normality

```{r}
taxa_nested_outliers_removed_lm_ln_plot_fitted_resid_qq <- taxa_nested_outliers_removed_lm_ln_plot_fitted_resid |>
  mutate(qq = map2(.x = lm_ln_object,
                  .y = taxa_clean,
                  .f = ~ ggplot(.x, aes(sample = residuals(.x))) + 
                    geom_qq(aes(sample = residuals(.x))) + 
                    stat_qq_line(aes(sample = residuals(.x))) +
                    ggtitle(.y) +
                    labs(x = "theoretical quantiles", y = "sample quantiles") +
                    theme_bw() + 
                    theme(panel.border = element_blank(), panel.grid.major = element_blank(), 
                    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"), 
                    axis.text = element_text(size = 12, colour = "black"), axis.title = element_text(size = 14),
                    title = element_text(size = 18))))
taxa_nested_outliers_removed_lm_ln_plot_fitted_resid_qq$qq
```

### Regression summary statistics

Pulling out regression summary statistics

```{r}
# Getting coefficients from lm
taxa_nested_outliers_removed_lm_ln_plot_coef <- taxa_nested_outliers_removed_lm_ln_plot |>
  mutate(summary = map(.x = lm_ln_object,
                       .f = ~ summary(.x)))
long_summary_outliers_removed <- taxa_nested_outliers_removed_lm_ln_plot_coef |>
  mutate(coefs = map(.x = summary,
                            ~.x[[4]]),
         r2 = map(.x = summary,
                            ~.x[[9]]),
         n = map(.x = lm_ln_object,
                 ~length(.x$residuals)),
         conf_interval = map(.x = lm_ln_object,
                             .f = ~confint(.x))) |>
  unnest(c(coefs, r2, n, conf_interval)) |>
  mutate(coef_type = c(rep(c("intercept", "slope"), times = length(taxa_clean)/2))) |>
  pivot_wider(names_from = coef_type, values_from = c(coefs, conf_interval))

intercept <- long_summary_outliers_removed$coefs_intercept[,1]
intercept_se <- long_summary_outliers_removed$coefs_intercept[,2]
intercept_ci_lower <- long_summary_outliers_removed$conf_interval_intercept[,1]
intercept_ci_upper <- long_summary_outliers_removed$conf_interval_intercept[,2]
slope <- long_summary_outliers_removed$coefs_slope[,1]
slope_se <- long_summary_outliers_removed$coefs_slope[,2]
slope_ci_lower <- long_summary_outliers_removed$conf_interval_slope[,1]
slope_ci_upper <- long_summary_outliers_removed$conf_interval_slope[,2]
p <- long_summary_outliers_removed$coefs_slope[,4] # p-value of the slope is the same as the p-vale of the overall regression model in the case of simple regression with one predictor
long_summary_outliers_removed_short <- long_summary_outliers_removed |>
  select(taxa_clean, n, r2)
  
reg_summary_outliers_removed <- add_column(long_summary_outliers_removed_short, 
                                           tibble(p, slope, slope_se, slope_ci_lower, slope_ci_upper, intercept, intercept_se, intercept_ci_lower, intercept_ci_upper))

# Back-transforming intercept to get parameter 'a' in power equation: y = ax^b
reg_summary_outliers_removed_a <- reg_summary_outliers_removed |> 
  mutate(a = exp(intercept),
         a_ci_lower = exp(intercept_ci_lower),
         a_ci_upper = exp(intercept_ci_upper))
```

Calculating SF (smearing factor - mean of back-transformed residuals) as per Mährlein, M., M. Pätzig, M. Brauns and A. M. Dolman. Length–mass relationships for lake macroinvertebrates corrected for back-transformation and preservation effects. Hydrobiologia, 768, 37-50 (2016).

```{r}
taxa_nested_outliers_removed_lm_ln_plot_coef_sf <- taxa_nested_outliers_removed_lm_ln_plot_coef |>
  mutate(sf = map(.x = lm_ln_object,
                  .f = ~ (1/length(residuals(.x)))*sum(exp(residuals(.x)))))

taxa_outliers_removed_sf <- taxa_nested_outliers_removed_lm_ln_plot_coef_sf |>
  select(taxa_clean, sf) |>
  unnest(sf)

# Adding smearing factor to lm summary table
reg_summary_outliers_removed_a_sf <- merge(reg_summary_outliers_removed_a, taxa_outliers_removed_sf)

# Saving the regression summary
write_csv(reg_summary_outliers_removed_a_sf,
          here("output/ethanol/dm_afdm/biomass_regression_summary_ethanol_dm_afdm_outliers_removed.csv"))
```

### Back-transformation 

Plotting back-transformed regression line with smearing factor correction and 95% confidence band

```{r}
# Extracting lower and upper 95% confidence bounds on the transformed scale
taxa_nested_outliers_removed_lm_ln_plot_coef_CI <- taxa_nested_outliers_removed_lm_ln_plot_coef_sf |> 
  mutate(conf_int = map(.x = lm_ln_object,
                        .f = ~ predict(.x, interval = "confidence"))) |>
  unnest(cols = c(data, sf, conf_int))

lower_CI <- taxa_nested_outliers_removed_lm_ln_plot_coef_CI$conf_int[,2]
upper_CI <- taxa_nested_outliers_removed_lm_ln_plot_coef_CI$conf_int[,3]

# Back-transforming confidence bounds and plotting with back-transformed regression line
taxa_nested_outliers_removed_lm_ln_plot_coef_CI_BT <- taxa_nested_outliers_removed_lm_ln_plot_coef_CI |>
  add_column(lower_CI, upper_CI) |>
  mutate(lower_CI_line = exp(lower_CI)*sf,
         upper_CI_line = exp(upper_CI)*sf) |>
  select(taxa_clean, dm_g, afdm_g, lower_CI_line, upper_CI_line, sf) |>
  nest(data = c(dm_g, afdm_g, lower_CI_line, upper_CI_line)) |>
  right_join(reg_summary_outliers_removed_a) |>
  mutate(BT_plot = pmap(.l = list(data, slope, intercept, sf, taxa_clean, r2),
                        .f = ~ggplot(..1, aes(x = dm_g, y = afdm_g)) +
                          geom_point(fill = "grey", size = 3, alpha = 0.5, stroke = 1) +
                          geom_function(fun = function(x) (exp(..3))*(x^..2)*..4, 
                                        col = "darkgreen", linewidth  = 1) +
                          geom_ribbon(aes(ymin = lower_CI_line, ymax = upper_CI_line), fill = "darkgreen", 
                                      alpha = 0.2) +
                          ggtitle(..5) + 
                          labs(x = "dry mass (g)", y = "ash-free dry mass (g)") + 
                          geom_text(x = 0.25*max(..1$dm_g), y = 0.85*max(..1$afdm_g), 
                                    label = deparse(bquote(paste("A = ", 
                                                                 .(format(round(exp(..3), digits = 2), nsmall = 2)),
                                                                 "D" ^.(format(round(..2, digits = 2), nsmall = 2))*
                                                                   " \u00D7 ",
                                                                 .(format(round(..4, digits = 2), nsmall = 2)), 
                                                                 sep = ""))),
                                    parse = TRUE, size = 4) +
                          geom_text(x = 0.25*max(..1$dm_g), y = 0.75*max(..1$afdm_g), 
                                    label = deparse(bquote(paste("R"^2, " = ", 
                                                                 .(format(round(..6, digits = 2), nsmall = 2)), 
                                                                 sep = ""))),
                                    parse = TRUE, size = 4) +
                          theme_bw() + 
                          theme(panel.border = element_blank(), 
                                panel.grid.major = element_blank(),
                                panel.grid.minor = element_blank(), 
                                axis.line = element_line(colour = "black"),
                                axis.text = element_text(size = 12, colour = "black"), 
                                axis.title = element_text(size = 14), 
                                title = element_text(size = 16)) + 
                          scale_x_continuous(expand = expansion(mult = c(0, 0.1)), limits = c(0, NA)) + 
                          scale_y_continuous(expand = expansion(mult = c(0, 0.1)), limits = c(0, NA))))
taxa_nested_outliers_removed_lm_ln_plot_coef_CI_BT$BT_plot
```

### Calculating biomass ratios and associated summary statistics

```{r}
biomass_ethanol_dm_afdm_outliers_removed_ratios <- biomass_ethanol_dm_afdm_cleaned |>
  filter(keep_data == TRUE) |>
  mutate(afdm_dm = afdm_g/dm_g) |>
  group_by(taxa_clean) |>
  summarise(n = n(),
            mean_afdm_dm = mean(afdm_dm),
            sd_afdm_dm = sd(afdm_dm),
            lower_ci_mean_afdm_dm = mean_afdm_dm-0.95*(sd_afdm_dm/sqrt(n)),
            upper_ci_mean_afdm_dm = mean_afdm_dm+0.95*(sd_afdm_dm/sqrt(n)),
            geom_mean = prod(afdm_dm)^(1/n),
            min_afdm_dm = min(afdm_dm),
            median_afdm_dm = median(afdm_dm),
            max_afdm_dm = max(afdm_dm),
            range_dm = paste0(format(round(min(dm_g), digits = 3), nsmall = 3), 
                              "-", format(round(max(dm_g), digits = 3), 
                                          nsmall = 3)))

# Saving the ratio summary
write_csv(biomass_ethanol_dm_afdm_outliers_removed_ratios,
          here("output/ethanol/dm_afdm/biomass_ratio_summary_ethanol_dm_afdm_outliers_removed.csv"))
```

### Assessing model performance

Calculating mean percentage error (MPE) and mean absolute percentage error (MAPE) for regression and ratio methods

```{r}
pred_mass_ethanol_dm_afdm_outliers_removed_error_eval <- taxa_nested_outliers_removed_lm_ln_plot_coef_CI_BT |>
  mutate(ols_pred_afdm = pmap(.l = list(data, a, slope, sf),
                            .f = ~ ..2*(..1$dm_g)^..3*..4),
         ols_pred_afdm_percent_error = map2(.x = data,
                                          .y = ols_pred_afdm,
                                          .f = ~-((.x$afdm_g-.y)/.x$afdm_g)*100),
         ratio_pred_afdm = map2(.x = data,
                              .y = biomass_ethanol_dm_afdm_outliers_removed_ratios$mean_afdm_dm,
                              .f = ~ .y*.x$dm_g),
         ratio_pred_afdm_percent_error = map2(.x = data,
                                          .y = ratio_pred_afdm,
                                          .f = ~-((.x$afdm_g-.y)/.x$afdm_g)*100),
         ols_total_PE = map2(.x = data,
                             .y = ols_pred_afdm,
                             .f = ~ (sum(.y-.x$afdm_g)/sum(.x$afdm_g))*100),
         ols_MPE = map2(.x = ols_pred_afdm_percent_error,
                        .y = n,
                        .f = ~ sum(.x)/.y),
         ols_MAPE = map2(.x = ols_pred_afdm_percent_error,
                       .y = n,
                       .f = ~ sum(abs(.x))/.y),
         ratio_total_PE = map2(.x = data,
                             .y = ratio_pred_afdm,
                             .f = ~ (sum(.y-.x$afdm_g)/sum(.x$afdm_g))*100),
         ratio_MPE = map2(.x = ratio_pred_afdm_percent_error,
                        .y = n,
                        .f = ~ sum(.x)/.y),
         ratio_MAPE = map2(.x = ratio_pred_afdm_percent_error,
                       .y = n,
                       .f = ~ sum(abs(.x))/.y)) |> 
  select(taxa_clean, ols_total_PE, ols_MPE, ols_MAPE, ratio_total_PE, ratio_MPE, ratio_MAPE) |> 
  unnest(c(ols_total_PE, ols_MPE, ols_MAPE, ratio_total_PE, ratio_MPE, ratio_MAPE))
```

### combining summary tables

```{r}
summary_ethanol_dm_afdm_ratios <- left_join(reg_summary_a_sf, 
                                         biomass_ethanol_dm_afdm_ratios) |>
  left_join(pred_mass_ethanol_dm_afdm_error_eval)

summary_ethanol_dm_afdm_ratios_outliers_removed <- left_join(reg_summary_outliers_removed_a_sf,
                                                          biomass_ethanol_dm_afdm_outliers_removed_ratios) |>
  left_join(pred_mass_ethanol_dm_afdm_outliers_removed_error_eval) |>
  mutate(taxa_clean = paste0(taxa_clean, "_outliers_removed"))

full_summary_ethanol_dm_afdm <- bind_rows(summary_ethanol_dm_afdm_ratios, summary_ethanol_dm_afdm_ratios_outliers_removed) |>
  arrange(taxa_clean)

write_csv(full_summary_ethanol_dm_afdm,
          here("output/ethanol/dm_afdm/full_summary_ethanol_dm_afdm.csv"))
```